
---
title: "Glaucoma Biomarker Discovery using DIABLO"
author: "tjmb03"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ️️ Goal

Identify multi-omics biomarkers associated with **primary open-angle glaucoma (POAG)** for potential diagnostic use.

##  Load Required Package

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("mixOmics")

library(mixOmics)
```

## 🧬 Simulate Multi-Omics Data

```{r}
set.seed(123)

n <- 100  # 50 POAG, 50 controls
Y <- factor(rep(c("POAG", "Control"), each = 50))

mRNA <- matrix(rnorm(n * 1000), nrow = n, ncol = 1000)
colnames(mRNA) <- paste0("gene_", 1:1000)

proteomics <- matrix(rnorm(n * 200), nrow = n, ncol = 200)
colnames(proteomics) <- paste0("prot_", 1:200)

metabolomics <- matrix(rnorm(n * 50), nrow = n, ncol = 50)
colnames(metabolomics) <- paste0("met_", 1:50)

X <- list(mRNA = mRNA, proteomics = proteomics, metabolomics = metabolomics)
```

## 🔧 Define Design and DIABLO Model

```{r}
design <- matrix(1, ncol = 3, nrow = 3)
diag(design) <- 0
colnames(design) <- rownames(design) <- names(X)

keepX <- list(mRNA = c(30, 30), proteomics = c(20, 20), metabolomics = c(10, 10))

diablo_model <- block.splsda(X, Y, ncomp = 2, keepX = keepX, design = design)
```

## 📈 Visualization

```{r, fig.width=10, fig.height=8}
plotIndiv(diablo_model, legend = TRUE, title = "Sample Clustering by DIABLO")
plotVar(diablo_model, var.names = TRUE, legend = TRUE)
circosPlot(diablo_model, cutoff = 0.3)
network(diablo_model, cutoff = 0.3)
```

## ✅ Performance Evaluation

```{r}
perf_diablo <- perf(diablo_model, validation = "Mfold", folds = 5, nrepeat = 10)
plot(perf_diablo)
```

##  Extract Biomarker Signatures

```{r}
selected <- selectVar(diablo_model, comp = 1)

top_genes <- selected$mRNA$name
top_proteins <- selected$proteomics$name
top_metabolites <- selected$metabolomics$name

top_genes
top_proteins
top_metabolites
```

## Next Steps (Suggestions)

- Run GSEA or pathway enrichment on selected genes.
- Use STRING or Cytoscape for PPI network analysis.
- Upload metabolite list to MetaboAnalyst for interpretation.
- Validate biomarker panel on independent dataset.
